---
output:
  html_document: default
  pdf_document: default
---
---
 title: "Integrative Bayes Model"
 author: "Shuang Jiang et al"
 date: "12/03/2018"
 output: rmarkdown::html_vignette
 vignette: >
   %\VignetteIndexEntry{Vignette Title}
   %\VignetteEngine{knitr::rmarkdown}
   %\VignetteEncoding{UTF-8}
 ---

 ```{r setup, include = FALSE}
 knitr::opts_chunk$set(
   collapse = TRUE,
   comment = "#>"
 )
 ```
# Tutorial

The following script is used to fit micribiome count data and covariate data to the 
integrative Bayesian zero-inflated negative binomial hierarchical mixture model 
proposed in the manuscript

Before running the following code, please first load micribiome count data and covariate data.
The necessary inputs should be

1. a n-by-p count matrix Y, where n is the number of samples and p is the number 
of taxa(feature)
2. a n-by-R covaritae matrix X, where R is the number of covariates
3. a n-dimensional vector z, indicating group allocation for n samples. Currently, we support the analysis of 2 phenotypes, and each element in vector z takes the value of 1 or 2.

You also need to install `Rcpp`, `RcppArmadillo` and `pROC` packages.

##  Load functions & data matrices
```{r load}
library(IntegrativeBayes)
load(system.file("extdata/Example_data.Rdata", package = "IntegrativeBayes"));
```

## Preprocessing
```{r preprocess}
# keep the features that have at least 2 observed counts for both groups:
Y.input = Y.filter(Y.mat, zvec = z.vec, min.number = 2)[[2]]
# estimate the size factor s from the count matrix Y:
s.input = sizefactor.estimator(Y.mat)
```

## Get true label for later visualization
```{r label}
feature.remain = which(Y.filter(Y.mat, zvec = z.vec, min.number = 2)[[1]] == 1)
gamma.vec = gamma.vec[feature.remain]
delta.mat = delta.mat[,feature.remain]
```

## Implement MCMC algorithm
```{r mcmc}
S.iter = 10000
burn.in = 0.5
res = zinb_w_cov(Y_mat = Y.input,
                 z_vec = z.vec, 
                 s_vec = s.input,
                 X_mat = X.mat,
                 S = S.iter, burn_rate = burn.in)
```

The MCMC outputs are stored in `res`:
    $mu0_posterior_mean: posterior mean(after burn-in) for the vector mu(0j)
    $beta_posterior_mean: posterior mean(after burn-in) for the Beta matrix
    $gamma_PPI: PPI for all gamma(j) after burn-in
    $delta_PPI: PPI for all delta(rj) fter burn-in
    $R_PPI: PPI for all r(ij) after burn-in
    $gamma_sum: sum of all gamma(j) for each iteration
    $mukj_store: MCMC draws for mu(kj) after burn-in (k = 1; j = 1,...,p)
    $mu0_store: MCMC draws for mu(0j) after burn-in (j = 1, ..., p)
    $beta_store: MCMC draws for beta(rj) after burn-in (r = 1, .., R; j = 1, ..., p )

## Visualizing the results for two variable selection processes

### Variable selection for discriminating features

The stem-plot showed the selected features passing Bayesian FDR threshold.

```{r visualization}
gamma_VS(res$gamma_PPI, gamma.true = gamma.vec)
par(mar=c(5.1, 4.1, 4.1, 2.1))
```

### Variable selection for significant feature-covariate association

The ROC plot was used to benchmark the performance of detecting microbiome-covariate associations.

```{r visualization2}
delta_ROC(as.vector(res$delta_PPI), as.vector(abs(delta.mat)))
```


# Citation

S. Jiang, G. Xiao, A. Y. Koh, Q. Li, and X. Zhan, "A Bayesian Zero-Inflated Negative Binomial Regression Model for the Integrative Analysis of Microbiome Data".

# Contact 

Shuang Jiang <shuangj@smu.edu>
Last updated on Mar 5, 2019.

